---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readxl)
library(tidyverse)
# For GO enrichment analysis
library(clusterProfiler)
# gene ids
library(org.Hs.eg.db)
```

## Exclusive

Loading the overview exclusive excel sheet of interest

```{r}
exclusive <- readxl::read_excel("overview_exclusive_table.xlsx",col_names = FALSE)
```

Removing some of the extra columns and naming some empty columns

```{r}
exclusive[1,1] <- "State"
exclusive[2,1] <- "Cancer"
exclusive[3,1] <- "Patient"
exclusive <- exclusive[,-c(101:102)]
exclusive <- as.data.frame(t(exclusive))
```

Summation of exclusive families in different datasets

```{r}
colnames(exclusive) <- exclusive[1,]
intermediate_aggregation <- exclusive %>% 
  filter(!row_number() %in% c(1, 2)) %>%
  dplyr::select(-c(Patient,Dataset)) %>%
  group_by(State,Cancer) %>%
  mutate_if(is.character, as.numeric) %>%
  summarise(across(where(is.numeric), sum))
```

Done like this because Head and Neck cancer metastasis are also included with the cancer patients, and are not excluded in getting the common families

```{r}
filtered <- intermediate_aggregation %>%
  filter(State %in% c("Tumour","Lymph node metastasis")) %>%
  group_by(Cancer) %>%
  dplyr::select(-State) %>%
  aggregate(.~Cancer,FUN = sum)

filtered[12,] <- c("Sum",filtered %>%  summarise(across(where(is.numeric), sum)))

filtered <- as.data.frame(t(filtered))
```

```{r}
colnames(filtered) <- filtered[1,]
filtered <- filtered[-1,]
```

First step is to determine the limit to which family enrichment can be considered

```{r}
selected_families <- colSums(sapply(filtered,as.numeric))
no_of_patients <- c(9,10,1,6,22,6,4,5,6,6,1,76)
names(no_of_patients) <- names(selected_families)
```

```{r}
p_cal <- function(no_of_patients,selected_families,i) {
  fraction <- selected_families/(1144*no_of_patients)
  p_for_i <- choose(no_of_patients,i)*(fraction**i)*((1-fraction)**(no_of_patients-i))
  return(p_for_i)
}

binomial_p <- function(sel_fam,pat_no,test_no) {
  i <- seq(test_no,pat_no,1)
  return(sum(p_cal(pat_no,sel_fam,i))*1144)
}

binomial_limit <- function(no_of_patients,selected_families) {
  if (no_of_patients == 1){
    # need to write it like this (instead of just 2) for cancer identity to show later
    return(no_of_patients+1)
  }
  test_no <- no_of_patients
  p <- 0
  while (p<0.05) {
    i <- seq(test_no,no_of_patients,1)
    p <- sum(p_cal(no_of_patients,selected_families,i))*1144
    test_no <- test_no - 1
  }
  # test_no + 2 because
  # test_no was reduced by 1 before the end of the while loop (+1)
  # test_no was non-significant, implying that test_no+1 was the number of interest for us (+1)
  return(test_no+2)
}
```

```{r}
filtered_intermediate <- as.data.frame(sapply(filtered,as.numeric))
rownames(filtered_intermediate) <- rownames(filtered)

exc_fam <- list()
for (i in 1:12){
 exc_fam[[i]] <- rownames(filtered_intermediate[filtered_intermediate[,i] >= binomial_limit(no_of_patients[i],selected_families[i]),])
}
```

```{r}
union <- filtered_intermediate[Reduce(union,exc_fam),]
write.csv(union,"exclusive_enriched_families_union_all_datasets.csv")
```

### Family name Heatmap

```{r}
panther <- read.csv("clean_pantherhuman.csv")
panther <- panther[,c(4,6)] %>% unique
colnames(panther)[which(names(panther) == "family_id")] <- "Families"
famnames <- panther[panther$Families %in% rownames(union),]
famnames <- famnames[match(rownames(union),famnames$Families),]
famnames$Families <- paste0("(", famnames$Families, ")")
famnames$Families <- gsub("PTHR","",famnames$Families)
famnames$combined <- paste0(famnames$family_name," ",famnames$Families)
row.names(union) <- famnames$combined
```

```{r}
library(ComplexHeatmap)
library(colorRamp2)
num_fam <- exclusive %>%
  group_by(State) %>%
  filter(State %in% c("Tumour","Lymph node metastasis")) %>%
  .$Cancer %>%
  table
relative_freq <- as.data.frame(t(t(union[,1:11])/as.vector(num_fam)))
relative_freq <- as.matrix(relative_freq)
col_fun = colorRamp2(c(0, 1), c("white", "red"))
Heatmap(relative_freq,
        column_names_max_height = unit(12,"cm"),
        row_names_max_width = unit(14,"cm"),
        name = "Proportion",
        col = col_fun)
```

```{r}
union <- filtered_intermediate[Reduce(union,exc_fam),]
union[union > 0] <- 1
union <- rowSums(union[-12])
exclusive_to_select <- names(union[union>2])
```

## GO

```{r}
panther <- read.csv("clean_pantherhuman.csv")
gene_id <- as.character(panther[panther$family_id %in% exclusive_to_select,]$gene_symbol)
```

getting rid of genes that have 0 expression in total even when all the 76 tumour patients are taken into account

```{r}
gene_id <- gene_id[! gene_id %in% c('OPN1MW3')]
```

```{r}
go <- enrichGO(gene_id, keyType = "SYMBOL",'org.Hs.eg.db', ont="ALL",minGSSize = 4,
  maxGSSize = 5000, pvalueCutoff=1, pAdjustMethod = "BH")
```

```{r}
go_df <- go@result
go_df$GeneRatioNum <- sapply(go_df$GeneRatio, function(x) eval(parse(text=x)))
go_df$BgRatioNum <- sapply(go_df$BgRatio, function(x) eval(parse(text=x)))
go_df$Gene_by_Bg <- go_df$GeneRatioNum/go_df$BgRatioNum
```

```{r}
go_df$geneID <- strsplit(as.character(go_df$geneID),"/")

for (i in 1:nrow(go_df)){
  tr <- as.data.frame(go_df$geneID[[i]])
  colnames(tr) <- "gene_symbol"
  contribution <- unique(merge(tr,panther)[,"family_id"])
  go_df$Families[i] <- paste(contribution, collapse = ", ")
  go_df$NumFamContribution[i] <- length(contribution)
}

for (i in 1:nrow(go_df)){
go_df$geneID[i] <- paste(go_df$geneID[[i]],collapse = ", ")
}
go_df$geneID <- unlist(go_df$geneID)

write.csv(go_df,"enrichGO_baseOutput_9Families_minusEmptyGene.csv")
```

```{r}
go_df$geneID <- strsplit(go_df$geneID,", ")
go_df <- go_df[(go_df$NumFamContribution>1) & (go_df$p.adjust<0.05) & (go_df$ONTOLOGY=="BP") & (go_df$Count > 20),]
```

```{r}
proportion_matrix <- function(df){
  sapply(1:length(df$geneID), 
         function(i1) sapply(1:length(df$geneID),function(i2) length(intersect(df$geneID[[i1]],df$geneID[[i2]]))/length(df$geneID[[i1]])) )
}
```

```{r}
mat_go_df <- proportion_matrix(go_df)
colnames(mat_go_df) <- go_df$Description
rownames(mat_go_df) <- go_df$Description
write.csv(mat_go_df,"Similarity_Matrix_BP_9Families.csv")
```

### Bubble Plot with \# of genes contributing to family

```{r}
# filtered go
select_terms <- c("transport","metal ion transport","transmembrane transport","multicellular organism development","cell differentiation","cell migration","regulation of cell population proliferation")
go_df <- go_df[match(select_terms,go_df$Description),]
```

```{r}
panther <- read.csv("clean_pantherhuman.csv")
for (i in 1:nrow(go_df)){
  tr <- as.data.frame(go_df$geneID[[i]])
  colnames(tr) <- "gene_symbol"
  contribution <- merge(tr,panther)[,"family_name"]
  go_df$FamiliesRedundant[i] <- paste(contribution, collapse = ", ")
}
```

```{r}
gene_num <- go_df %>% separate_rows(FamiliesRedundant, sep = ", ")
```

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num,
       aes(x = stringr::str_wrap(str_to_sentence(FamiliesRedundant),20), 
           y = fct_rev(fct_inorder(stringr::str_wrap(str_to_sentence(Description),20))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom") +
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(panel.background = element_rect(colour = "black", fill=NA, size=1),
        panel.border = element_blank(),
        legend.position = "right",
        axis.text.x = element_text(size = 23, vjust = 0.5, hjust=0.5, angle = 45),
        axis.text.y = element_text(size = 30, vjust = 0.5, hjust=0.5,face = c('plain', 'plain','bold','bold','plain','plain','bold')),
        plot.title = element_text(size=20, vjust = 0.5, hjust=0.5),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title=element_text(size=20), 
        legend.text=element_text(size=18),
        legend.key = element_rect(fill = "transparent", colour = "transparent"))
# element_rect(colour = "black", fill=NA, size=1)
```

### Alternate Bubble Plots

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num,
       aes(y = str_to_sentence(FamiliesRedundant),
           x = fct_rev(fct_inorder((stringr::str_wrap(str_to_sentence(Description),20)))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom") +
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(panel.background = element_rect(colour = "black", fill=NA, size=1),
        panel.border = element_blank(),
        legend.position = "right",
        axis.text.y = element_text(size = 30, vjust = 0.5, hjust=0.5),
        axis.text.x = element_text(size = 22, vjust = 0.5, hjust=0.5,face = c('plain', 'plain','bold','bold','plain','plain','bold'), angle = 45),
        plot.title = element_text(size=20, vjust = 0.5, hjust=0.5),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title=element_text(size=15), 
        legend.text=element_text(size=14),
        legend.key = element_rect(fill = "transparent", colour = "transparent"),
        axis.ticks = element_blank())
```

```{r}
go_df_alt <- go_df[,1:15]
for (i in 1:nrow(go_df_alt)){
  tr <- as.data.frame(go_df_alt$geneID[[i]])
  colnames(tr) <- "gene_symbol"
  contribution <- merge(tr,panther)[,"family_id"]
  go_df_alt$FamiliesRedundant[i] <- paste(contribution, collapse = ", ")
}
gene_num_alt <- go_df_alt %>% separate_rows(FamiliesRedundant, sep = ", ")
```

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num_alt,
       aes(x = stringr::str_wrap(FamiliesRedundant,20), 
           y = fct_rev(fct_inorder(str_to_sentence(Description))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom") +
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(panel.background = element_rect(colour = "black", fill=NA, size=1),
        panel.border = element_blank(),
        legend.position = "right",
        axis.text.x = element_text(size = 28,vjust = 0.5, hjust=0.5,angle = 45),
        axis.text.y = element_text(size = 30, vjust = 0.5, hjust=0.5,face = c('plain', 'plain','bold','bold','plain','plain','bold')),
        plot.title = element_text(size=20, vjust = 0.5, hjust=0.5),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.key = element_rect(fill = "transparent", colour = "transparent"),
        axis.ticks = element_blank())
```

```{r}
panther <- read.csv("clean_pantherhuman.csv")
for (i in 1:nrow(go_df)){
  tr <- as.data.frame(go_df$geneID[[i]])
  colnames(tr) <- "gene_symbol"
  contribution <- merge(tr,panther)[,"family_name"]
  go_df$FamiliesRedundant[i] <- paste(contribution, collapse = ", ")
}
gene_num <- go_df %>% separate_rows(FamiliesRedundant, sep = ", ")
```

```{r}
gene_num_abbrev <- gene_num
gene_num_abbrev[gene_num_abbrev == "MULTIPLE EPIDERMAL GROWTH FACTOR-LIKE DOMAINS PROTEIN"] <- "MEGF-like domains protein"
gene_num_abbrev[gene_num_abbrev == "FIBROBLAST GROWTH FACTOR"] <- "Fibroblast growth factor"
gene_num_abbrev[gene_num_abbrev == "MYELIN TRANSCRIPTION FACTOR 1-RELATED"] <- "MyT1-related"
gene_num_abbrev[gene_num_abbrev == "CALCIUM-ACTIVATED POTASSIUM CHANNEL SUBUNIT BETA"] <- bquote("K[Ca] subunit β")
gene_num_abbrev[gene_num_abbrev == "SMALL CONDUCTANCE CALCIUM-ACTIVATED POTASSIUM CHANNEL"] <- "small conductance K[Ca]"
gene_num_abbrev[gene_num_abbrev == "T-CELL LYMPHOMA BREAKPOINT-ASSOCIATED TARGET 1-RELATED"] <- "TCBA target 1-related"
gene_num_abbrev[gene_num_abbrev == "SOLUTE CARRIER FAMILY 5"] <- "SLC5"
gene_num_abbrev[gene_num_abbrev == "SODIUM/POTASSIUM/CALCIUM EXCHANGER"] <- "Na/K/Ca Exchanger"
gene_num_abbrev[gene_num_abbrev == "OPSIN"] <- "Opsin"

gene_num_abbrev <- gene_num_abbrev %>% mutate(FamiliesRedundant=gsub(" ", "~", FamiliesRedundant))
```

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num_abbrev,
       aes(x = str_to_sentence(FamiliesRedundant), 
           y = fct_rev(fct_inorder(stringr::str_wrap(str_to_sentence(Description),20))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom",labels=parse(text=sort(unique(gene_num_abbrev$FamiliesRedundant))))+ # Because of this, it is no longer 
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(panel.background = element_rect(colour = "black", fill=NA, size=1),
        panel.border = element_blank(),
        legend.position = "right",
        axis.text.x = element_text(size = 30, vjust = 0.5, hjust=0.5, angle=45),
        axis.text.y = element_text(size = 30, vjust = 0.5, hjust=0.5,face = c('plain', 'plain','bold','bold','plain','plain','bold')),
        plot.title = element_text(size=20, vjust = 0.5, hjust=0.5),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.key = element_rect(fill = "transparent", colour = "transparent"),
        axis.ticks = element_blank())
```

#### With shortened descriptions for two GO terms

```{r}
gene_num[gene_num == "multicellular organism development"] <- "Multi. organism dev."
gene_num[gene_num == "regulation of cell population proliferation"] <- "Reg. of cell pop. proliferation"

gene_num_alt[gene_num_alt == "multicellular organism development"] <- "Multi. organism dev."
gene_num_alt[gene_num_alt == "regulation of cell population proliferation"] <- "Reg. of cell pop. proliferation"

gene_num_abbrev[gene_num_abbrev == "multicellular organism development"] <- "Multi. organism dev."
gene_num_abbrev[gene_num_abbrev == "regulation of cell population proliferation"] <- "Reg. of cell pop. proliferation"
```

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num,
       aes(y = FamiliesRedundant,
           x = fct_rev(fct_inorder(str_to_sentence((stringr::str_wrap(Description,20))))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom") +
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(legend.position = "right",
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 20,face = c('plain', 'plain','bold','bold','plain','plain','bold'), angle = 45),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank())
```

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num_alt,
       aes(x = stringr::str_wrap(FamiliesRedundant,20), 
           y = fct_rev(fct_inorder(str_to_sentence(Description))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom") +
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(legend.position = "right",
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 24,face = c('plain', 'plain','bold','bold','plain','plain','bold')), ,
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank())
```

```{r}
cols2 <- c(rep("A",98),rep("B",29),rep("C",68))
ggplot(gene_num_abbrev,
       aes(x = stringr::str_wrap(FamiliesRedundant,20), 
           y = fct_rev(fct_inorder(str_to_sentence(Description))),colour = cols2))+
  geom_count() +  
  guides(color="none") + 
  scale_colour_manual(values =c("blue","pink","red"))+
  scale_x_discrete(position = "bottom",labels=parse(text=sort(unique(gene_num_abbrev$FamiliesRedundant))))+ # Because of this, it is no longer 
  scale_size_continuous(range = c(5, 30)) + # Adjust as required.
  labs(x = NULL, y = NULL,size = "No. of genes", colour = "Super Category") +
  theme(legend.position = "right",
        axis.text.x = element_text(size = 18, angle = 45),
        axis.text.y = element_text(size = 24,face = c('plain', 'plain','bold','bold','plain','plain','bold')), ,
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank())
```
